How to obtain the redshift distribution from probabilistic redshift estimates

### 2018-06-28 research/paper/draft.tex
significant progress

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\textul}{\underline}
\newcommand{\chippr}{\texttt{CHIPPR} }

\shorttitle{How to obtain the redshift distribution from probabilistic redshift 
estimates}
\shortauthors{Malz, et al.}

@@

\title{How to obtain the redshift distribution from probabilistic redshift 
estimates}

\author{Alex Malz\altaffilmark{1}, David W. Hogg\altaffilmark{1,2,3,4}, Phil 
Marshall\altaffilmark{5}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
Physics,
  New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
Heidelberg, Germany}
\altaffiltext{5}{[SLAC]}

\begin{abstract}
A trustworthy estimate of the redshift distribution $n(z)$ is crucial for weak 
lensing cosmology.  Spectroscopically confirmed redshifts for the dim and 
numerous galaxies of weak lensing surveys are expected to be inaccessible, 
making photometric redshifts (photo-$z$s) the next best alternative.  The 
nontrivial systematics affecting photo-$z$ estimation have motivated the weak 
lensing community to favor photo-$z$ probability distribution functions (PDFs) 
as a more comprehensive alternative to photo-$z$ point estimates.  However, 
analytic methods for utilizing these new data products in cosmological 
inference are still evolving.  This paper presents the Cosmological 
Hierarchical Inference with Probabilistic Photometric Redshifts (CHIPPR) model, 
a novel approach to estimating the $n(z)$ from a catalog of galaxy photo-$z$ 
PDFs based upon a probabilistic graphical model of hierarchical inference.  We 
present the publicly available \chippr code implementing this technique, as 
well as its validation on mock data.  The CHIPPR model yields a more accurate 
characterization of $n(z)$ containing information beyond the best-fit estimator 
produced by traditional procedures.

\end{abstract}

@@

@@

After a brief literature review addressing how photo-$z$ PDFs are currently 
used in cosmology, this paper aims to answer the following questions:

\begin{itemize}
	\item Why should we question existing methods?
	\item How can we improve the effectiveness of using photo-$z$ PDFs in 
inference?
	\item How does the result of \chippr compare to established estimators 
in terms of the accuracy of $n(z)$?
	\item How significant is the effect of the discrepancy between $n(z)$ 
estimators on cosmological constraints?
\end{itemize}

\section{Method}
\label{sec:method}

This paper presents a mathematically consistent method for obtaining the 
posterior probability distribution over the redshift distribution $n(z)$ using 
a catalog of photo-$z$ PDFs.  We start by introducing some nomenclature, 
definitions, and symbols that will be used throughout Secs. \ref{sec:model}, 
\ref{sec:others}, and \ref{sec:implementation}.

We shall say the parameters comprising $\vec{\phi}$ define the redshift 
distribution under some functional form about which this method is agnostic, so 
long as the functional form evaluated with some true values of the parameters 
accurately describes the true $n(z)$.  Since the redshift distribution is 
itself a probability distribution, it may be written as $p(z | \vec{\phi})$.

Before jumping into the details of the model, it is crucial to settle on an 
interpretation of what a photo-$z$ PDF actually is.  When a photo-$z$ PDF is 
reported, it is an object containing information about the redshift $z_{i}$ of 
galaxy $i$ based on its photometric data $\vec{d}_{i}$, making a photo-$z$ PDF 
a \textit{posterior} probability distribution.  The data $\vec{d}_{i}$ may be 
fluxes, magnitudes, colors, or any combination thereof.  We assume that the 
redshifts of the galaxies are not related to the redshifts or photometry of 
other galaxies in the survey, i.e. they are \textit{independent}.

In addition to the data $\vec{d}_{i}$, the photo-$z$ PDF also contains 
information imparted by assumptions that went into the process by which the 
photo-$z$ was made.  Those assumptions are only relevant insomuch as they 
impose a preference on the redshift distribution. Thus, they can be reduced to 
a sort of prior on the distribution of redshifts, which can be parametrized 
under the chosen functional form by parameters $\vec{\phi}^{*}$.  The redshift 
distribution $p(z | \vec{\phi}^{*})$ associated with the parameters in 
$\vec{\phi}^{*}$ shall be called the interim prior because in many cases it was 
our best guess as to the true redshift distribution before observing any data.  
For example, in template-based photo-$z$ PDF methods, the interim prior is a 
linear combination of the redshift distributions of the classes of templates 
based on previous observations.  (In training-based photo-$z$ PDF methods, the 
interim prior is related to the redshift distribution of the training set and 
thus may not be shared among all galaxies for which photo-$z$ PDFs are 
produced; however, in this work, we will assume that it is shared among all 
galaxies in a survey.)  Because the choice of parameters $\vec{\phi}^{*}$ is 
not causally connected to the data or the true redshift distribution but 
nonetheless contributes to the photo-$z$ PDF, we call it an \textit{interim 
prior}.

A photo-$z$ PDF is then an \textit{interim posterior} probability distribution 
of the redshift of a galaxy given its observed photometric data and the interim 
prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  
We present the two attributes of a probabilistic graphical model (PGM) for the 
redshift distribution $n(z)$: the directed acyclic graph (DAG) and its 
mathematical interpretation in terms of Bayesian hierarchical inference.

The fundamental assumption underlying the concept of photo-$z$ estimation is 
that each galaxy $i$ has some observed photometric data $\vec{d}_{i}$ (fluxes, 
magnitudes, or colors) that is drawn from a function of its redshift $z_{i}$, 
which is a parameter in this model.  This function constitutes a forward model 
for the observations.  The redshifts $\{z_{i}\}$ for all galaxies in a survey 
are random draws from the redshift distribution, whose parameters  $\vec{\phi}$ 
under a chosen functional form are called \textit{hyperparameters} because they 
are shared among all $N$ galaxies in the survey.  Fig. \ref{fig:pgm} 
illustrates these relationships.  PGMs of this structure are 
\textit{hierarchical} in that, though the data are only directly influenced by 
their redshift parameters, we can still use them to infer something about the 
global hyperparameters by way of a higher level in the graph.

\begin{figure}
	\begin{center}We would like to learn about $p(z | \vec{\phi})$ from a 
catalog of photo-$z$ interim posteriors.

		\subsection{Model}
		\label{sec:model}

		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
		\caption{This directed acyclic graph corresponds to a 
probabilistic graphical model for a Bayesian hierarchical inference of 
$p(\vec{\phi}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown 
in circles, with observed random variables shown in shaded circles.  
Relationships between variables are indicated by arrows from parameters to the 
variables distributed according to functions of them.  The box indicates that 
there are a number of copies of the relationships between boxed parameters, 
each independent of all others.  The hyperparameters comprising $\vec{\phi}$, 
which define $n(z)$, are at the top.  Independently drawn from a function of 
the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.  The 
observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is 
determined by the redshifts above.}
	\label{fig:pgm}
	\end{center}
\end{figure}

The DAG of Fig. \ref{fig:pgm} translates directly into a mathematical 
expression for the posterior distribution $p(\vec{\phi} | \{\vec{d}_{i}\})$ of 
the hyperparameters given the entire set of interim posteriors $\{p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})\}$ by way of the following derivation.  From this 
point forward, we will use log probabilities.

According to Bayes' rule, the posterior distribution of interest is
\begin{align}
\label{eq:hyper_bayes}
\log[p(\vec{\phi} | \{\vec{d}_{i}\})] &\propto \log[p(\{\vec{d}_{i}\} | 
\vec{\phi})] + \log[p(\vec{\phi})] ,
\end{align}
where $p(\vec{\phi})$ is the \textit{hyperprior} probability distribution over 
possible values of the hyperparameters.  The hyperprior is a choice made by 
those performing the inference and is not to be confused with the interim prior 
parameters that influence the photo-$z$ PDF data product.

Focusing on the likelihood term $p(\{\vec{d}_{i}\} | \vec{\phi})$, we employ 
the power of the hierarchical model through \textit{marginalization} of the 
unobserved redshift parameters by integrating over them to obtain 
\begin{align}
\label{eq:marginalize}
\begin{split}
\log[p(\{\vec{d}_{i}\} | \vec{\phi})] = & 
\log\left[\int\exp\left[\log[p(\{\vec{d}_{i}\} | \{z_{i}\})]\right.\right. \\
& \left.\left.+ \log[p(\{z_{i}\} | \vec{\phi})]\right]\ d\{z_{i}\}\right] .
\end{split}
\end{align}
The redshifts $\{z_{i}\}$ are statistically independent of each other, and each 
galaxy's photometry depends solely on its own redshift, so we may invoke 
statistical independence to get
\begin{align}
\label{eq:independence_hyper}
\log[p(\{z_{i}\} | \vec{\phi})] &= \sum_{i}^{N} \log[p(z_{i} | \vec{\phi})]
\end{align}
and
\begin{align}
\label{eq:independence}
\log[p(\{\vec{d}_{i}\} | \{z_{i}\})] &= \sum_{i}^{N} \log[p(\vec{d}_{i} | 
z_{i})] .
\end{align}
Having chosen the functional form mapping $\vec{\phi}$ to $n(z)$, the terms on 
the righthand side of Eq. \ref{eq:independence_hyper} are known.  However, the 
terms on the righthand side of Eq. \ref{eq:independence} are likelihoods, and 
the photo-$z$ PDFs we have are interim posteriors.

To transform the interim posteriors we have into the likelihoods we need, we 
start with the vacuously true statement
\begin{align}
\begin{split}
\label{eq:unity}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
\end{split}
\end{align}
We expand the last term using Bayes' Rule to get
\begin{align}
\begin{split}
\label{eq:bayes}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and 
$\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ 
as follows:
\begin{align}
\begin{split}
\label{eq:split}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(\vec{d}_{i} | z_{i})] - \log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Canceling terms, we find
\begin{align}
\label{eq:cancel}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - 
\log[p(z_{i} | \vec{\phi}^{*})].
\end{align}

Now we combine Eqs. \ref{eq:hyper_bayes}, \ref{eq:marginalize}, 
\ref{eq:independence_hyper}, \ref{eq:independence}, and \ref{eq:cancel} to 
arrive at 
\begin{align}
\begin{split}
\label{eq:final}
\log[p(\vec{\phi} | \{\vec{d}_{i}\})] \propto& \log[p(\vec{\phi})] + 
\log\left[\int \exp\left[\sum_{i}^{N} \left(\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | \vec{\phi})] - \log[p(z_{i} | 
\vec{\phi}^{*})]\right.\right.\right.\\
& \left.\left.\left. + \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})]
\right)\right] d\{z_{i}\}\right].
\end{split}
\end{align}
Thus, if we have a catalog of photo-$z$ interim posteriors $\{p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})\}$, the interim prior parameters $\vec{\phi}^{*}$, 
and a prior distribution $p(\vec{\phi})$, we may find the posterior 
$\log[p(\vec{\phi} | \{\vec{d}_{i}\})]$ on the redshift distribution $n(z)$.

This framework entails a number of choices and assumptions that must be 
addressed explicitly:
\begin{enumerate}
	\item The chosen functional form of $n(z)$ with parameters $\vec{\phi}$ 
must be capable of describing the true redshift distribution.
	\item The photometric data and redshift for each galaxy must be 
independent of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior shared among all galaxies in the 
survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
	\item We must choose a hyperprior probability distribution 
$p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the 
probabilistic graphical model presented here is not the only one that could be 
proposed.
\end{enumerate}

\subsection{Alternative Approaches}
\label{sec:others}

It is useful to translate some popular existing methods for deriving $n(z)$ 
from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.  
We briefly discuss the conditions under which they are equivalent to Eq. 
\ref{eq:final}, which are elaborated upon in the Appendix.

\subsubsection{Stacking}
\label{sec:stacking}

The most common way to combine photo-$z$ interim posteriors into an estimator 
$\hat{n}_{stack}(z)$ for $n(z)$ is to "stack" them, which corresponds to
\begin{align}
\label{eq:stack}
\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} \exp\left[\log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\right].
\end{align}

\subsubsection{Point Estimation}
\label{sec:point_estimation}

In some cases, estimators $\hat{n}^{point}(z)$ are obtained by assuming the 
photo-$z$ PDFs are effectively delta functions by reducing them to point 
estimates according to
\begin{align}
\label{eq:delta}
p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) &\approx \delta(z, \hat{z}_{i}^{point})
\end{align}
before applying Eq. \ref{eq:stack}.  The most popular redshift point estimators 
are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode 
$\hat{z}^{mode}_{i}$ of the original photo-$z$ interim posterior.

@@

The publicly available \chippr code implements the probabilistic graphical 
model presented in Sec. \ref{sec:model}.

In addition to the choices and assumptions underlying the probabilistic 
graphical model, the implementation of \chippr makes choices and assumptions of 
its own.

\begin{enumerate}
	\item \chippr is only applicable if the redshift interim posteriors and 
interim prior are accurate.
	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
samples of a single format, that of the piecewise constant parametrization, 
also referred to as a binned histogram parametrization and a sum of top hat 
functions.
\end{enumerate}

\section{Validation on Simple Mock Data}
\label{sec:validation}

We demonstrate the superiority of \chippr over alternative approaches in a 
number of compelling test cases on mock data.  Each experiment is characterized 
by a single change to a fiducial case in order to isolate the influence of 
systematic effects known to be relevant to photo-$z$ estimation and propagation 
in analysis.

\subsection{Mock Data \& Metrics}
\label{sec:validintro}

The mock data in these tests consists of photo-$z$ interim posteriors rather 
than photometric data because the various existing methods for deriving 
photo-$z$ interim posteriors do not in general yield results that are 
consistent with one another, indicating that their systematics are not 
well-understood.  Because the mock data is independent of any choice of 
photo-$z$ PDF production method, we not only ensure that our photo-$z$ interim 
priors are perfectly understood but also deter readers from assuming that 
\chippr has any preference over the method by which photo-$z$ interim 
posteriors are derived from photometric data.

\subsubsection{Mock Data}
\label{sec:mockdata}

The mock data used here are produced by the following steps.

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
function with known parameters $\vec{\phi}'$.
	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
$N$ true redshifts $z'_{i}$.
	\item \label{itm:scatter} Choose a true intrinsic scatter parameter 
$\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = 
\mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining 
Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, 
\sigma)$.
	\item \label{itm:interim} Choose a parametrization and the parameters 
$\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | 
\vec{\phi}^{*})]$ in this parametrization.
\end{enumerate}

In all of the following validation tests, we use a piecewise constant 
parametrization in log-space with $10$ bins.  $N = 10,000$ galaxies.  In Secs. 
\ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.  This 
method for deriving mock data is referred to as the fiducial case, and 
variations on it will refer directly to the steps that are altered.

\subsubsection{$n(z)$ Accuracy Metric}
\label{sec:accuracy}

The Kullback-Leibler divergence is our primary measure of the accuracy of 
estimators of $n(z)$ in cases of mock data with known true redshifts.

\textbf{Review precision and bias from \texttt{kld.ipynb} and interpret in 
terms of a \% difference.}

\subsection{Underlying $n(z)$ Effects}
\label{sec:truth}

Existing $n(z)$ estimators are systematically smoother than the true $n(z)$.  
Here we show that the traditional estimators perform better when the true 
$n(z)$ is weakly featured than when the true $n(z)$ is strongly featured by 
experimenting with Step \ref{itm:true} of Sec. \ref{sec:mockdata}.  The 
implication of this issue is quite serious; the consistently smooth, unimodal 
$n(z)$ estimates appearing in the literature could result from much more 
featured true redshift distributions, and there would be no way to catch this 
error without using a fully probabilistic method.

\subsubsection{Featureless $n(z)$}
\label{sec:smooth}

In this test, we choose a weakly featured true $n(z)$ of Fig. \ref{fig:smooth} 
with a smooth, unimodal shape, based on the interim prior used for the SDSS DR8 
photo-$z$ PDFs.  [check and include citation!]

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/smooth_truth/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/smooth_truth/estimators.png}	
		\caption{All estimators perform well when the true $n(z)$ is 
well behaved, exhibiting significant deviation only when $n(z)$ is very small, 
as the sample size of true redshifts in that range will be small.  Top panel: 
The traditional MAP reduction of a photo-$z$ PDF against the true redshifts 
with a few rescaled photo-$z$ interim posteriors are overplotted in solid 
lines, with a dotted line indicating zero probability.  Bottom panel: Various 
estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ as a 
continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

\subsubsection{Featured $n(z)$}
\label{sec:featured}

In this test, we choose the true $n(z)$ of Fig. \ref{fig:featured} with 
nontrivial structure.

\begin{figure}
	\begin{center}

\includegraphics[width=0.5\textwidth]{fig/fiducial/scatter.png}\\

\includegraphics[width=0.5\textwidth]{fig/fiducial/estimators.png}		
		\caption{[This plot doesn't really show what I want because the 
intrinsic scatter is too low!  I used $\sigma=0.03$ because that's what has 
been quoted as what LSST, etc. needs, but that's not what comes out of 
photo-$z$ estimation methods before they impose aggressive cuts\dots]  Top 
panel: The traditional MAP reduction of a photo-$z$ PDF against the true 
redshifts with a few rescaled photo-$z$ interim posteriors are overplotted in 
solid lines, with a dotted line indicating zero probability.  Bottom panel: 
Various estimators of $\ln[n(z)]$, the interim prior, and the true $\ln[n(z)]$ 
as a continuous function and under a binned parametrization.}
		\label{fig:featured}
	\end{center}
\end{figure}

This featured true $n(z)$ will henceforth be referred to as the fiducial $n(z)$

\subsection{Emulated Data Quality Effects}
\label{sec:likelihoods}

In the following test cases, we vary the properties of the mock photo-$z$ 
likelihoods in an effort to emulate known systematics in photo-$z$ estimation.  
These tests vary Step \ref{itm:scatter} of Sec. \ref{sec:mockdata}.

\subsubsection{Intrinsic Scatter}
\label{sec:intscat}

>>>>

\usepackage{soul}
\usepackage[title]{appendix}

\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}{\underline}
\newcommand{\chippr}{CHIPPR}
\newcommand{\nz}{$n(z)$}
\newcommand{\pz}{photo-$z$}
\newcommand{\pzpdf}{photo-$z$ PDF}

@@

\title{How to obtain the redshift distribution from probabilistic redshift 
estimates}%:\\

\author{Alex Malz\altaffilmark{1}}
\author{David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of 
Physics,
  New York University, 726 Broadway, 9th floor, New York, NY 10003, USA}
\altaffiltext{2}{Simons Center for Computational Astrophysics, 162 Fifth 
Avenue, 7th floor, New York, NY 10010, USA}
\altaffiltext{3}{Center for Data Science, New York University, 60 Fifth Avenue, 
7th floor, New York, NY 10003, USA}
\altaffiltext{4}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 
Heidelberg, Germany}

\begin{abstract}
A trustworthy estimate of the redshift distribution $n(z)$ is crucial for 
weak-lensing cosmology as we know it.
Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing 
surveys are expected to be inaccessible, making photometric redshifts 
(photo-$z$s) the next-best alternative.
The nontrivial systematics affecting photo-$z$ estimation have motivated the 
weak-lensing community to favor photo-$z$ probability density functions (PDFs) 
as a more comprehensive alternative to photo-$z$ point estimates.
However, analytic methods for utilizing these new data products in cosmological 
inference are still evolving.
The ubiquitous methodology known as stacking produces a systematically biased 
estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very 
regime where photo-$z$ PDFs are most necessary.
We introduce a mathematically rigorous probabilistic graphical model (PGM) of 
hierarchical inference of $n(z)$, which is provably the only self-consistent 
way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
The novel Cosmological Hierarchical Inference with Probabilistic Photometric 
Redshifts (CHIPPR) model yields a more accurate characterization of $n(z)$ by 
correctly propagating the redshift uncertainty information beyond the best-fit 
estimator produced by traditional procedures.
We conclude by propagating these effects to constraints in the space of 
cosmological parameters.
\end{abstract}

@@

\maketitle

@@

\textcolor{blue}{Brief literature review addressing how \pzpdf s are currently 
used in cosmology}

\textcolor{blue}{Q: Why should we question existing methods?\\
A: Stacking is bad and only looks like it works because of assumptions that 
don't hold when the data is as bad as we anticipate.  Cite pedantic doc?}

overview of next questions: How can we improve the effectiveness of using 
\pzpdf s in inference?  How does the result of \chippr\ compare to established 
estimators in terms of the accuracy of \nz ?  \textcolor{blue}{Reach goal: How 
significant is the effect of the discrepancy between \nz\ estimators on 
cosmological constraints?}

\section{Methods}
\label{sec:methods}

\textcolor{blue}{Q: How can we improve the effectiveness of using \pzpdf s in 
inference?\\
A: Hierarchical inference is the only self-consistent way.}

This paper presents the Cosmological Hierarchical Inference with Probabilistic 
Photometric Redshifts (\chippr ) model as a mathematically consistent method 
for constraining \nz s using a catalog of \pzpdf s.
In Section~\ref{sec:model}, we introduce the \chippr\ formalism; in 
Section~\ref{sec:application}, we outline how to use \chippr\ to constrain \nz\ 
from a catalog of \pzpdf s; in Section~\ref{sec:limitations} we document the 
assumptions that must hold for \chippr\ to be valid.

\subsection{Model generalities}
\label{sec:model}

We begin by reframing the redshift distribution \nz\ from a probabilistic 
perspective.
Here we define a redshift distribution \nz\ as the normalized probability 
density
\begin{equation}
\label{eqn:nz}
\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ 
\int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
\end{equation}
of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
We may without loss of generality impose a parametrization
\begin{equation}
\label{eqn:fz}
f_{\phi}(z)\ \equiv\ n(z)
\end{equation}
in terms of some parameter vector $\phi$.
We believe that galaxy redshifts are indeed drawn from \nz, making it a 
probability density over redshift; this fact can also be confirmed from 
dimensional analysis of Equation \ref{eqn:nz}.
Therefore, it can be rewritten as
\begin{equation}
\label{eqn:pz}
z_{j}\ \sim\ \Pr(z \mid \phi)\ \equiv\ f_{\phi}(z),
\end{equation}
a probability density over redshift conditioned on the parameters defining \nz.
Note that $z_{j}$ does not depend on $z_{j'}$, a statement of the causal 
independence of galaxy redshifts from one another.

In addition to believing \nz\ is a PDF from which redshifts are drawn, we also 
believe that there is some PDF from which photometric data $d$, which may be 
any combination of fluxes, magnitudes, colors, and their observational errors, 
are drawn.
Such a PDF over data is a likelihood
\begin{equation}
\label{eqn:pzpdf}
d\ \sim\ \Pr(d \mid z)
\end{equation}
conditioned on redshift.
This assumption that the data are drawn from some function of the redshift 
forms the foundation upon which \pz\ estimation is based.
Note that galaxies may have different observational data $d$ despite sharing 
the same redshift and that the data $d_{j}$ of one galaxy is causally 
independent of the redshifts $z_{j'}$ and data $d_{j'}$ of other galaxies.

This description of the physical system corresponds to a forward model by which 
we actually believe photometry is generated:
\begin{enumerate}
	\item There exists a redshift distribution \nz\ with parameters $\phi$.
	\item Galaxy redshifts $\{z_{j}\}$ are independent draws from $\Pr(z 
\mid \phi)$.
	\item Galaxy photometry $d_{j}$ is drawn from the likelihoods $\Pr(d | 
z_{j})$.
\end{enumerate}
A forward model such as this corresponds to a probabilistic graphical model 
(PGM), represented by a directed acyclic graph (DAG) as in Figure~\ref{fig:pgm}.
\begin{figure}
	\begin{center}
		\includegraphics[width=0.25\textwidth]{fig/pgm.png}
		\caption{The PGM}
		\label{fig:pgm}
	\end{center}
\end{figure}
A DAG conveys the causal relationships between physical parameters and, like a 
Feynman diagram in the context of particle physics, is a shorthand for 
mathematical relationships between variables.

\subsection{Application to cosmology}
\label{sec:application}

The problem facing cosmologists is to determine the true parameters $\phi_{0}$ 
of \nz\ from observing the photometry $\{d_{j}\}$ of a large sample of galaxies 
$j$.
To self-consistently propagate the uncertainty in the redshifts, however, it is 
more appropriate to estimate the posterior $\Pr(\phi \mid \{d_{j}\})$ over all 
possible parameters $\phi$ (and thus potential redshift distributions \nz ) 
conditioned on all the observed data $\{d_{j}\}$ available in a generic catalog.

In order to use the DAG of Figure~\ref{fig:pgm} to derive an expression for 
$\Pr(\phi \mid \{d_{j}\})$ in terms of \pzpdf s, we must introduce two more 
concepts, confusingly named the implicit prior and the prior probability 
density.

When we constrain the redshift of a galaxy using its observed photometric data 
$d_{j}$, we are effectively estimating a posterior $\Pr(z \mid d_{j})$.
However, to do this, we must have a model for the general relationship between 
redshifts and photometry, whether empirical, as is the case for machine 
learning \pzpdf\ methods, or analytic, as is the case for template-based 
\pzpdf\ methods.
Such a relationship is defined in the space of probability density over 
redshift, so it must be able to be parameterized by the same functional form 
$f_{\phi}(z)$ as \nz .
It is thus natural to write it as $\Pr(z \mid \phi^{*})$, where $\phi^{*}$ is 
the parameters for this relationship under some generic \pzpdf\ method.
We call $\Pr(z \mid \phi^{*})$ the \textit{implicit prior}, as it is rarely 
explicitly known nor chosen by the researcher; for template-based methods where 
it can be chosen to be ``realistic,'' it may be appropriate to call it an 
\textit{interim prior}.
Because the implicit prior is unavoidable, the \pzpdf s reported by any method 
are really \textit{implicit-weighted posteriors} $\Pr(z \mid d, \phi^{*})$.

The prior probability density $\Pr(\phi)$ is a more familiar concept in 
astronomy; to progress, we will have to choose a prior probability density over 
all possible parameters $\phi$.
This prior need not be excessively proscriptive; for example, it may be chosen 
to enforce smoothness at physically motivated scales in redshift without 
imposing any particular region as over- or under-dense.

With these definitions, we obtain the desired expression for $\Pr(\phi \mid 
\{d_{j}\})$,
\begin{align}
\begin{split}
  \label{eqn:fullpost}
  \ln[\Pr(\phi \mid \{d_{j}\})] & \propto \ln[\Pr(\phi)] + \ln \left[\int dz 
\right.\\
  & \left. \exp \left[\sum_{j=1}^{J} \left(\ln[\Pr(z \mid d_{j}, \phi^{*})] 
\right. \right. \right.\\
  & \left. \left. \left. + \ln[\Pr(z \mid \phi)] - \ln[\Pr(z \mid \phi^{*})]\ 
\right)\ \right]\ \right] ,
\end{split}
\end{align}
which is the heart of \chippr.
The entire derivation of Equation~\ref{eqn:fullpost} is provided in 
Appendix~\ref{app:math}.

@@

In this study, we compare the results of Equation~\ref{eqn:fullpost} to those 
of the two most common approaches to estimating \nz\ from a catalog of \pzpdf 
s: the distribution $n(z_{\mathrm{max}})$ of the redshifts at maximum posterior 
probability (i.e. the modes of the \pzpdf s), and the stacked estimator
\begin{align}
  \label{eqn:stacked}
  \hat{n}(z) &\equiv \frac{1}{J}\ \sum_{j=1}^{J}\ \Pr(z \mid d_{j}, \phi^{*}) .
\end{align}

We perform this comparison on mock data in the form of catalogs of emulated 
\pzpdf s generated via the forward model discussed above and presented in 
detail in Appendix~\ref{app:data}.
The mock data emulates the three sources of error of highest concern to the 
\pz\ community: intrinsic scatter, catastrophic outliers, and systematic bias.
Figure~\ref{fig:mega_scatter} illustrates these three effects individually at 
ten times the tolerance of the upcoming Large Synoptic Survey Telescope (LSST).
\begin{figure*}
	\begin{center}

\includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
    \includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
    \includegraphics[width=0.3\textwidth]{fig/forward_model_example.png}
		\caption{Show this for each individually.}
		\label{fig:mega_scatter}
	\end{center}
\end{figure*}
Tests including all three effects at the tolerance levels of LSST and at twice 
the tolerance levels of LSST are presented in Section~\ref{sec:results}.

\subsection{Limitations}
\label{sec:limitations}

Finally, we explicitly review the assumptions made by this approach.
\begin{itemize}
  \item prior $\Pr(\phi)$
  \item implicit prior $\phi^{*}$ known
	\item \pzpdf s are accurate posteriors %draws from photo-$z$ PDFs do 
not match marginal distributions in $z_true$ vs. $z_phot$ space, i.e. toy model 
with $z_phot$ drawn from an uglier space but pdfs evaluated in lsst-limits space
	\end{itemize}

\section{Results}
\label{sec:results}

\textcolor{blue}{Q: How does the result of \chippr compare to established 
estimators in terms of the accuracy of $n(z)$?\\
A: \chippr yields the best possible $n(z)$, conditional on the accuracy of the 
photo-$z$ PDFs used.}

\subsection{Idealized example}
\label{sec:idealized}

\textcolor{blue}{LSST specs: unbiased, uniform outliers, medium z-dependent 
scatter}

\begin{figure*}
	\begin{center}

\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
		\caption{Show this for idealized case.}
		\label{fig:idealized}
	\end{center}
\end{figure*}

\subsection{Realistic example}
\label{sec:realistic}

\textcolor{blue}{DES(?) specs: biased to enhance true $n(z)$, systematic 
outliers, high z-dependent scatter}

\begin{figure*}
	\begin{center}

\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
		\caption{Show this for realistic case.}
		\label{fig:realistic}
	\end{center}
\end{figure*}

\subsection{Violations of the model}
\label{sec:violations}

\textcolor{blue}{mischaracterized interim prior, with idealized or realistic 
data?}

\begin{figure*}
	\begin{center}

\includegraphics[width=0.45\textwidth]{fig/forward_model_example.png}
		\includegraphics[width=0.45\textwidth]{fig/results_example.png}
		\caption{Show this for mischaracterized case.}
		\label{fig:mischaracterized}
	\end{center}
\end{figure*}
